---
title: "Tidying Data in R"
author: "Vera Kuklina"
date: "8/15/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



packageName::functionName
example:

stats::filter()

```{r, warning = F, message = F}
library(dplyr)
library(tidyr)
```
Read in data from Mike Byerly. 2016. Alaska commercial salmon catches by management region (1886- 1997). urn:node:GOA. df35b.304.2.


```{r}
catch_df <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1", method = "libcurl"),
                    stringsAsFactors = FALSE)
head(catch_df)
```
Move from wide to long table

```{r}
catch_df <- catch_df %>% 
  select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum)
# or select (-All, -notesRegCode)
```


```{r}
catch_df <- catch_df %>% 
  gather(key = "species", value = "catch", -Region, -Year) %>% 
  rename(catch_thousands = catch)

head(catch_df)
```

Reverse of gather is spread

```{r}
catch_integers <- as.integer(catch_df$catch_thousands)

which(is.na(catch_integers) == T)
```


1) turn "I" into 1 in the catch_thousands column
2) coerce catch_thousands into integer type
3) convert catch_thousands to catch

```{r}
catch_df <- catch_df %>% 
  mutate(catch_thousands = ifelse(catch_thousands == "I", 1, catch_thousands)) %>% 
  mutate(catch_thousands = as.integer(catch_thousands)) %>% 
  mutate(catch = catch_thousands * 1000)

tail(catch_df)
```

# Split-Apply-Combine

```{r}
catch_summarized <- catch_df %>%
  group_by(Region) %>% 
  summarize(mean_catch = mean(catch))

head(catch_summarized)
  
```

```{r}
catch_range <- catch_df %>% 
  group_by(Region) %>% 
  summarize(catch_low = range(catch)[1], 
            catch_high = range(catch)[2])

head(catch_range)
```


```{r}
catch_mean <- catch_df %>% 
  group_by(Region, species) %>% 
  #comment - can write whatever I want. Hear I calculate mean, standard deviation and number of rows
  summarize(catch_mean = mean(catch, na.rm = T),
            catch_stdev = sd(catch),
            num_obs = n()) %>% 
  #filter for Chinook
  filter(species == "Chinook")

print(catch_mean)
```

```{r}
region_defs <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1", method = "libcurl"),
                        stringsAsFactors = FALSE)
```

# Joins
Read in region definition files
```{r}
region_defs <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1", method = "libcurl"),
                        stringsAsFactors = FALSE)

head(region_defs)
```

```{r}
region_defs <- region_defs %>% 
  select(code, mgmtArea)

head(region_defs)
```

```{r}
catch_joined <- left_join(catch_df, region_defs, by = c("Region" = "code"))

head(catch_joined)
```

